Efficient Feature Screening for Lasso-Type Problems via Batched Safe-Strong Rules

### 2019-12-16 main.tex
Modified last part of section 1 and beginning of section 2.

One promising technique to address this challenge is feature screening. The lasso model solution is sparse in the sense that most coefficients will be exactly zero; for these ``inactive'' coefficients we do not need their associated features.  In theory, if we knew which coefficients were nonzero, we could read into memory only those features and leave the remaining features on disk. In reality, we cannot know this prior to fitting the model -- however, through the clever use of feature screening rules, we can greatly reduce the number of times inactive features are read into memory, thereby minimizing the computational burden.  In what follows, we refer to features left on disk as ``discarded'' features, although it is worth clarifying that this is not a permanent decision: for example, a feature can be discarded at one value of $\lambda$ but included in the list of potentially active features at other values of $\lambda$.

There are two important categories of features screening rules: ``strong'' rules\cite{tibshirani2011regression}, which occasionally incorrectly discard an active feature, and ``safe'' rules\cite{ghaoui2010safe,wang2013lasso,xiang2012fast, xiang2011learning}, which are guaranteed never to do so.  As one might expect, strong rules tend to be easy to calculate; however, because mistakes are possible, one must add a post-convergence check to ensure that no features were incorrectly discarded.  This check is not necessary with safe rules.  However, safe rules are considerably less powerful at discarding features than strong rules.  More recently, hybrid approaches \cite{zeng2017efficient} combining the two types of rules have been shown to be more efficient than either type of rule alone.

\note{I think we need a better transition here} In this paper, we proposed adaptive batched safe-strong rules. We divided the set of $\lambda$ values into batches and screening within each batch is based on the same solution at one $\lambda$ value at the beginning of the batch. On one hand, basing on the same solution avoids many costly computation and on the other hand, referring to solution at a close $\lambda$ value provides more screening power. An adaptive choice of batches sizes are applied using the screening performance from previous $\lambda$ values to better balance these two effects. It will be shown that our adaptive rule can greatly reduce the computation time and outperform other method under all scenarios. Similar to HSSR, our method can also utilize any member of safe rules, giving it the potential for further speed up given new safe rules and the flexibility to extend to other lasso type models. For example we extended the method to work with logistic regression with lasso penalty.

@@

where $r(\lambda)$ is the residuals vector at $\lambda$ defined as $y-X\hat{\beta}(\lambda)$. If the residuals are known, we can safely conclude feature $j$ has 0 coefficient if the second condition meets, but residuals cannot be calculated without first solving the coefficients. However, when fitting a solution path, residuals at previous $\lambda$ values can be calculated and provide information about residuals at current $\lambda$ value.

@@

\noindent In this algorithm, residuals $r$ are computed at each $\lambda$ value for the update of the coefficients and can also be reused conveniently for feature screening purpose.

Before introducing current screening rules, there is also a similar idea to reduce the number of features for part of computation, called active-set cycling\cite{lee2007efficient}. When solving the coefficients for current $\lambda$, first consider the model with coefficients that were non-zero at previous $\lambda_0$. After computing the first solution, a post-convergence KKT conditions check is performed and if there is no violation, the solution to the smaller model will be the exact solution to the full model. If there are violations, corresponding features will be added to the model and the procedure will repeat. Because features don't enter the model very frequently, usually small number of repetitions will be needed and thus the algorithm converges fast.

@@

\subsection{}

>>>>

One promising technique to address this challenge is feature screening. The lasso model solution is sparse in the sense that most coefficients will be exactly zero; for these ``inactive'' coefficients we do not need their associated features.  In theory, if we knew which coefficients were nonzero at each value of $\lambda$, we could read into memory only those features and leave the remaining features on disk. In reality, we cannot know this prior to fitting the model -- however, through the clever use of feature screening rules, we can greatly reduce the number of times inactive features are read into memory, thereby minimizing the computational burden.  In what follows, we refer to features left on disk as ``discarded'' features, although it is worth clarifying that this is not a permanent decision and a decision is made for each $\lambda$ values: for example, a feature can be discarded at one value of $\lambda$ but included in the list of potentially active features at other values of $\lambda$.

There are two important categories of features screening rules: ``strong'' rules\cite{tibshirani2011regression}, which occasionally incorrectly discard some active features, and ``safe'' rules\cite{ghaoui2010safe,wang2013lasso,xiang2012fast, xiang2011learning}, which are guaranteed never to do so.  As one might expect, strong rules tend to be easier to calculate; however, because mistakes are possible, one must add a post-convergence check to ensure that no features were incorrectly discarded. This check is not necessary with safe rules. However, safe rules are considerably less powerful at discarding features than strong rules. More recently, hybrid approaches \cite{zeng2017efficient} combining the two types of rules have been shown to be more efficient than either type of rule alone.

\note{I think we need a better transition here} In this paper, we proposed an adaptive batched screening scheme that changes how screening is done along the path of $\lambda$ values and implement it into hybrid approaches to form adaptive batched safe-strong rules. We divided the set of $\lambda$ values into batches and screening within each batch is based on the same solution at one $\lambda$ value at the beginning of the batch. On one hand, basing on the same solution avoids many costly computation and on the other hand, referring to solution at a close $\lambda$ value provides more screening power. An adaptive choice of batches sizes are applied using the screening performance from previous $\lambda$ values to better balance these two effects. It will be shown that our adaptive rule can greatly reduce the computation time and outperform other method under all scenarios. Similar to hybrid approaches, our method can also utilize any member of safe rules, giving it the potential for further speed up given new safe rules and the flexibility to extend to other lasso type models. For example we extended the method to work with logistic regression with lasso penalty.

@@

where $r(\lambda)$ is the residuals vector at $\lambda$ defined as $y-X\hat{\beta}(\lambda)$. If the residuals are known, we can safely conclude feature $j$ has 0 coefficient if the second condition holds with "$<$", but residuals cannot be calculated without first solving the coefficients. However, when fitting a solution path, residuals at previous $\lambda$ values can be calculated and provide information about residuals at current $\lambda$ value and thus be used for screening.

@@

\noindent In this algorithm, residuals $r$ are computed at each $\lambda$ value for the update of the coefficients and can thus be reused conveniently for feature screening purpose.

@@

It will also be assumed that lasso model is solved along $K+1$ decreasing values of $\lambda$ denoted as $\lambda_0,\lambda_1,...,\lambda_K$.

\subsection{Active-set Cycling}

Before introducing current screening rules, there is also a similar idea to reduce the number of features for part of computation, called active-set cycling (AC) \cite{lee2007efficient}. When solving the coefficients for at $\lambda_k$, first consider the model with coefficients that were non-zero at $\lambda_{k-1}$. After computing the first solution, a post-convergence KKT conditions check is performed and if there is no violation, the solution to the smaller model will be the exact solution to the full model. If there are violations, corresponding features will be added to the model and the procedure will repeat. Because features don't enter the model very frequently, usually small number of repetitions will be needed and thus the algorithm converges fast.
